Preregistration: Changes

### 2022-07-25 manuscript.Rmd
- corrected typos

During the COVID-19 pandemic, 
numerous events unfolded in quick succession and several open questions emerged.
How dangerous is the virus? 
Is it spreading in my region? 
How is it transmitted, and how can I protect myself?
Because for many it was (and at the time of writing still is) a matter of life or death, people aimed to stay informed regarding the latest developments.
Governments around the world implemented safety measures, such as wearing masks, keeping physical distance, or enforcing lockdowns.
In this extraordinary situation, many people heavily relied on media to attain relevant information, and especially social media were at an all time high [@statistaAverageDailyTime2021].

@@

**In this study, to measure the effects of social media use focused on COVID-19 related news and topics, I adopt both the channel and the type of communication perspective.**
**Together, this should offer a nuanced and comprehensive understanding of communication.**
First, I investigate how well-being is affected by different types of communication affect, namely active and passive use.
Defining what constitutes active and what passive use is not always clear, and different understandings are currently discussed [@ellisonWhyWeDon2020; @meierComputermediatedCommunicationSocial2020a].
Reading is generally considered as passive and writing as active, while there are also specific behaviors such as such liking or sharing content that fall somewhere in-between [@meierDoesPassiveSocial2022].
In this study, I hence distinguish (a) _reading_ (passive), (b) _posting_ (active), and (c) _liking and sharing_ COVID-19 related posts (both active and passive).
Second, I analyze how using the most prominent branded applications affects well-being, and whether this effect changes across applications.
The branded applications investigated here are Facebook, Twitter, Instagram, WhatsApp, and YouTube---which were, at the time of writing, the most relevant social media apps.

@@

But people can also misjudge media affects and are often overly optimistic [@metzgerComparativeOptimismPrivacy2017].
And precisely because social media have so many positive consequences, one can ask if this is not where the actual problem lies.
In other words, social media aren't problematic because they're inherently bad, but rather because they're too good.
And as with many other things, there can be too much of a good thing.
It is therefore often asked whether social media are addictive, and users sometimes express this fear themselves [@yangCanWatchingOnline2021]. 
However, a recently published meta-analysis found that the two most prominent measures of addiction, the Bergen Facebook Addiction Scale and the Bergen Social Media Addiction Scale, have only small relations to well-being [@duradoniWellbeingSocialMedia2020]. 
In addition, the general idea of labeling excessive social and new media use as addiction was criticized, arguing that social media represent new regular behaviors that should not be pathologized [@galerHowMuchToo2018; @vanrooijWeakScientificBasis2018].

@@

There could also be _indirect_ effects.
When browsing social media for Covid-19 related news, many users reported being captivated to such an extent that they could not stop using social media [@kleinDarklySoothingCompulsion2021].
During the pandemic social media use was at an all-time high in the US [@statistaAverageDailyTime2021].
Although it is most likely that moderate social media use is not detrimental [@orbenTeenagersScreensSocial2020], overuse, however, might be more critical, and several studies have shown more pronounced negative effects for extreme users [@przybylskiLargescaleTestGoldilocks2017].
To explain, overuse could impair well-being if it replaces meaningful or functional activities such as meeting others, working, actively relaxing, or exercising.
**Another potentially negative mechanism at play are problematic social comparison processes. During the pandemic, several users shared how they successfully dealt with challenges such as physical distancing. In a study with 1131 residents from Wuhan in China [@yuePassiveSocialMedia2022], people who spent more time in quarantine also spent more time on social media. Those, who spent more time on social media also engaged in more upward social comparison, which was related to increased levels of stress.**
<!-- Therefore, if a society collectively overuses social media during a pandemic, there is potential for negative effects. -->
<!-- If overuse replaces such activities it's reasonable to assume that it's also detrimental. -->

@@

Finally, also several individual studies employing advanced methods found smalls relations between social media use and well-being [@orbenSocialMediaEnduring2019; @przybylskiDoesTakingShort2021a; @schemerImpactInternetSocial2021; @kerestesAdolescentsOnlineSocial2020].
For example, @beyensSocialMediaUse2021 reported that although for some users (roughly one quarter) the effects of social media use on well-being were negative, for almost the same number of users they were positive, while for the rest the effects were neutral.
This finding is aligned the Differential Susceptibility to Media Effects Model: 
Although there is substantial _variation_ of media effects for individual users, the _average_ effects reported in the literature are often small [@valkenburgDifferentialSusceptibilityMedia2013].

@@

What could be a minimally interesting, nontrivial effect?
Because this is a normative and ultimately philosophical question, there can never be a clear, single, or unanimous answer. 
<!-- In the end, it is a personal question. -->
However, it is still necessary and helpful to try provide such a plausible benchmark. 
I therefore suggest the following SESOI for this research question:

@@

Thus, only when controlling for _all_ relevant confounders, we can correctly estimate causality without bias [@rohrerThinkingClearlyCorrelations2018].
Obviously, we can never be entirely sure to have included all confounders, which makes absolute statements regarding causality virtually impossible.
In addition, when determining the overall causal effect, we need to make sure _not_ to control for mediating variables [@rohrerThinkingClearlyCorrelations2018], for doing so would bias our assessment of the causal effect.
Complicating matters further, it is often unclear if a variable is a mediator or a confounder.[^collider]
However, despite all these caveats, when controlling for relevant variables (that aren't mediators), we can be much more certain that we measured causality correctly.
The aim should therefore be to collect as many varying and non-varying confounders as possible (which I believe is seldom done in our field), while knowing that absolute certainty regarding causality cannot be reached.

@@

Finally, one precondition of causality is temporal order. 
The cause needs to precede the effect.
Finding the right interval between cause and effect is crucial.
For example, if we want to understand the effect of alcohol consumption on driving performance, it makes a big difference if driving performance is measured one minute, one hour, one day, or one week after consumption.
<!-- Finding the right interval is difficult. -->
If variables are stable, longer intervals are needed; if they fluctuate, shorter intervals.
In the case of well-being, we need shorter intervals for **the more fluctuating** positive and negative affect and longer ones for **the more stable** life satisfaction [@dienlinImpactDigitalTechnology2020].
**Using social media can have instant effects on mood [@marcianoDynamicsAdolescentsSmartphone2022].**
**Effects on life satisfaction often take longer to manifest, for example because media use leads to actual changes in specific behaviors, which then affect life satisfaction [@dienlinDisplacementReinforcementReciprocity2017].**
Choosing the right interval is challenging, because especially short intervals are hard to implement in practice, often requiring advanced methods such as experience sampling (also known as in situ measurement or ambulant assessment) [@schnauber-stockmannMobileDevicesTools2020].
In this study, I hence analyze how using social media during the last week affected positive and negative affect during the same week.
In other words, if people during the last week engaged in more COVID-19 related social media use than the usually do, did they feel better or worse during that week than they usually do?
Regarding life satisfaction, I implemented a longer interval.
If people during the last week used COVID-19 related social media more than the usually do, were they at the end of the week more or less satisfied with their lives than they usually are?
I hence analyze if when a person changes their social media diet, are there (a) _simultaneous_ changes in their affect and (b) _subsequent_ changes in their life satisfaction?
**These relations will be controlled for varying confounders, which fosters a causal interpretation.**
Similar approaches were implemented by other studies [@johannesNoEffectDifferent2022; @scharkowHowSocialNetwork2020], and they are considered a best practice approach toward analyzing causality [@bellFixedRandomEffects2019].

@@

Because the data were analyzed post-hoc, no a-priori sample size planning on the basis of power analyses was conducted.
The sample is large, and it is hence well-equipped to reliably detect also small effects.
<!-- , which is why no exact post hoc power analysis were conducted. -->
In addition, because such large samples easily generate significant _p_-values even for very small effects, it helps that the hypotheses were tested with a smallest effect size of interest-approach.
To this end, I adopted the interval testing approach as proposed by @dienesUsingBayesGet2014.
On the basis of the SESOI, I then defined a null region.
In what follows, I explain how I determined the SESOI and the null region.

@@

The factorial validity of the scales were tested with confirmatory factor analyses (CFA).
Because Mardia’s test showed that the assumption of multivariate normality was violated, I used the more robust Satorra-Bentler scaled and mean-adjusted test statistic (MLM) as estimator.
To avoid over-fitting, I tested the scales on more liberal fit criteria (CFI > .90, TLI > .90, RMSEA <. .10, SRMR < .10) [@klinePrinciplesPracticeStructural2016].
<!-- Finally, REWB-models cannot model latent variables. -->
<!-- To increase precision, I therefore exported factor scores from the CFAs for positive and negative affect. -->
Mean scores were used for positive and negative affect.
<!-- Respondents who answered less than 50% of all questions were removed. -->
**Missing responses were imputed using multiple imputation with predictive mean matching (five iterations, five data-sets).**
**Predictive mean matching allowed to impute also categorical variables.**
**All variables were imputed except the media use measures, as they were not collected on each wave.**
**All variables included in the analyses presented here were used to impute missing data.**
**For the main analyses, results were pooled across all five data-sets.**

@@

**There was a small and statistically nonsignificant trend that reading COVID-19 related content slightly increased life satisfaction (_b_ = `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(conf.high) %>% round(2)`], _p_ = `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(p.value) %>% my_round("p")`).**
**At the same time, there was also a small and statistically nonsignificant trend that reading COVID-19 related content decreased positive affect (_b_ = `r dat_fig_results_within %>% filter(dv == "Positive affect" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_within %>% filter(dv == "Positive affect" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_within %>% filter(dv == "Positive affect" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(conf.high) %>% round(2)`], _p_ = `r dat_fig_results_within %>% filter(dv == "Positive affect" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(p.value) %>% my_round("p")`).**

@@

For between-person relations, no a-priori hypotheses were formulated.
Results showed that no relation crossed or was completely outside of the SESOI.
Four relations were statistically significant.
**Respondents who across all waves used social media more frequently than others to read about COVID-19 related posts reported slightly lower levels of positive affect than others (_b_ `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between" & Analysis == "Regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`]).**
**Respondents who across all waves used social media more frequently than others to write COVID-19 related posts reported sligthly higher levels of negative affect than others (_b_ `r dat_fig_results_between %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_between %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_between %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`]).**
**At the same time, respondents who across all waves used social media more frequently than others to write COVID-19 related posts also reported slightly higher levels of positive affect (_b_ `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`]).**
**Finally, respondents who across all waves used YouTube more frequently than others also reported slightly higher levels of life satisfaction than others (_b_ `r dat_fig_results_between %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between" & Analysis == "Regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_between %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_between %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`]).**
<!-- However, note that the effect were still completely inside of the null region, hence not large enough to be considered practically relevant. -->

@@

**That said, several preliminary and subtle trends can be observed.**
**First, overall the results do suggest that effects of COVID-19 related social media use on well-being rather tend to take place in the negative as opposed to the positive spectrum.**
**For example, people who wrote more COVID-19 related posts than usual reported slightly lower levels of life satisfaction than usual.**
**Similarly, people who wrote more COVID-19 related posts than usual also reported slightly more negative affect.**
**When writing posts and comments on social media, people explicitly and more deeply engage with COVID-19 related content.**
**Unfortunately, in general online tonality is often more extreme, negative, or aggressive, which potentially affects their authors.**
**Note that I controlled for whether or not participants had an corona infection during a specific wave, which rules out the potential explanation that having an infection was the root cause of increased communication and reduced well-being.**

**The potential explanation that tonality might be a relevant factor at play here is also supported by the second trend.**
**People, who during the pandemic spend more time on Instagram than usual, also experienced less negative affect than usual.**
**Instagram is well-known for its positivity bias [@waterlooNormsOnlineExpressions2018].**
**Content is generally post positive, uplifting, and (self-)flattering.**
**It seems the often-criticized positivity bias on Instagram might have been somewhat beneficial in times of the pandemic.**
**The critique that the positivity bias necessarily leads to envy and negative feelings is one-sided, because positive content can also inspire and motivate users [@meierInstagramInspirationHow2020], which could be especially helpful in times of lockdown and home-office.**
**To provide a concrete example, Instagram was successfully used as an interactive communication channel for first year students to have a better start into their new degree, effectively complementing alternative learning platform tools [@yeTurningInformationDissipation2020].**

@@

**The results showed that it makes sense to analyze different communication types and communication channels, and that active and passive communication showed different results.**
**Liking and sharing content did not show any within-person effects. **
**Such rather low-key active behaviors do not seem to affect well-being at all.**
**Regarding passive use, reading COVID-19 related posts is more ambivalent; results showed some weak trends towards a positive effect on life satisfaction, but a negative affect on mood.**
**It might be that reading and informing oneself about COVID-19 on social media might be helpful in the long term, but more negative for short-term affect.**
**Finally, proactively engaging via writing posts, the most active form of communication analyzed here, showed only negative effects on well-being.**
**Together, the results support the findings from @valkenburgAssociationsActivePassive2022, who could not confirm the claim that active use is good, versus passive is bad.**
**Focusing on communication channels, YouTube seems to be more negative, whereas Instagram is likely more positive.**
**Again, these are only very small effects.**
**Future research might elaborate on these specific relations to probe their stability and relevance.**

**From a more political and societal perspective, I believe the results imply that it does makes sense critically to reflect upon COVID-19 related social media use.** 
**In terms of concrete recommendations, on average it might be slightly beneficial to post less actively about COVID-19 on social media and to spent less time on YouTube.**
**Potential positive effects, however, will for many users likely not be noticeable.**
**Results allow to reject a positive effect: Writing more posts on social media will likely not increase well-being.**
**At all events, engaging in COVID 19-related social media use should, on average, not be a cause for concern.**

@@

The current study analyzed whether changes in media use were related to changes in well-being, while controlling for several potential confounders. 
Together, this allowed for an improved perspective on assessing causality.
**However, the opposite effect is also plausible, namely that well-being affects media use [@zillmannMoodManagementCommunication1988].**
**While controlling for potential confounders can support claims of causality, the procedure implemented here cannot prove causality.**
Causality necessitates temporal order, and the cause needs to precede the effect.
**The challenge is that** regarding media use, such effects often happen immediately or shortly after use, necessitating intervals in the hours, minutes, or even seconds.
In many cases only experience sampling studies asking users in the very moment can produce such knowledge.
However, even then we don't know for certain if we actually measured the right interval.
Effects depend on the intensity of use or the length of the interval.
To borrow the words from @rohrerTheseAreNot2021, there is no such thing as "the" effect of social media use on well-being.
Hence, to document how effects unfold, future research needs to employ different study designs probing different time lags.
In addition, more thought needs to be invested in what relevant stable and varying factors we should include as control variables, and I hope this study provides a first step into this direction.

>>>>

During the COVID-19 pandemic, 
numerous events unfolded in quick succession and several open questions emerged.
How dangerous is the virus? 
Is it spreading in my region? 
How is it transmitted, and how can I protect myself?
Because for many it was (and at the time of writing still is) a matter of life or death, people aimed to stay informed regarding the latest developments.
Governments around the world implemented safety measures, such as wearing masks, keeping physical distance, or enforcing lockdowns.
In this extraordinary situation, many people heavily relied on media to obtain relevant information, and especially social media were at an all time high [@statistaAverageDailyTime2021].

@@

**In this study, to measure the effects of social media use focused on COVID-19 related news and topics, I adopt both the channel and the type of communication perspective.**
**Together, this should offer a nuanced and comprehensive understanding of communication.**
First, I investigate how well-being is affected by different types of communication, namely active and passive use.
Defining what constitutes active and what passive use is not always clear, and different understandings are currently discussed [@ellisonWhyWeDon2020; @meierComputermediatedCommunicationSocial2020a].
Reading is generally considered as passive and writing as active, while there are also specific behaviors such as liking or sharing content that fall somewhere in-between [@meierDoesPassiveSocial2022].
In this study, I hence distinguish (a) _reading_ (passive), (b) _posting_ (active), and (c) _liking and sharing_ COVID-19 related posts (both active and passive).
Second, I analyze how using the most prominent branded applications affects well-being, and whether this effect changes across applications.
The branded applications investigated here are Facebook, Twitter, Instagram, WhatsApp, and YouTube---which were, at the time of writing, the most relevant social media apps.

@@

But people can also misjudge media effects and are often overly optimistic [@metzgerComparativeOptimismPrivacy2017].
And precisely because social media have so many positive consequences, one can ask if this is not where the actual problem lies.
In other words, social media aren't problematic because they're inherently bad, but rather because they're too good.
And as with many other things, there can be too much of a good thing.
It is therefore often asked whether social media are addictive, and users sometimes express this fear themselves [@yangCanWatchingOnline2021]. 
However, a recently published meta-analysis found that the two most prominent measures of addiction, the Bergen Facebook Addiction Scale and the Bergen Social Media Addiction Scale, have only small relations to well-being [@duradoniWellbeingSocialMedia2020]. 
In addition, the general idea of labeling excessive social and new media use as addiction was criticized, arguing that social media represent new regular behaviors that should not be pathologized [@galerHowMuchToo2018; @vanrooijWeakScientificBasis2018].

@@

There could also be _indirect_ effects.
When browsing social media for COVID-19 related news, many users reported being captivated to such an extent that they could not stop using social media [@kleinDarklySoothingCompulsion2021].
During the pandemic social media use was at an all-time high in the US [@statistaAverageDailyTime2021].
Although it is most likely that moderate social media use is not detrimental [@orbenTeenagersScreensSocial2020], overuse, however, might be more critical, and several studies have shown more pronounced negative effects for extreme users [@przybylskiLargescaleTestGoldilocks2017].
To explain, overuse could impair well-being if it replaces meaningful or functional activities such as meeting others, working, actively relaxing, or exercising.
**Another potentially negative mechanism at play are problematic social comparison processes. During the pandemic, several users shared how they successfully dealt with challenges such as physical distancing. In a study with 1131 residents from Wuhan in China [@yuePassiveSocialMedia2022], people who spent more time in quarantine also spent more time on social media. Those, who spent more time on social media also engaged in more upward social comparison, which was related to increased levels of stress.**
<!-- Therefore, if a society collectively overuses social media during a pandemic, there is potential for negative effects. -->
<!-- If overuse replaces such activities it's reasonable to assume that it's also detrimental. -->

@@

Finally, also several individual studies employing advanced methods found smalls relations between social media use and well-being [@orbenSocialMediaEnduring2019; @przybylskiDoesTakingShort2021a; @schemerImpactInternetSocial2021; @kerestesAdolescentsOnlineSocial2020].
For example, @beyensSocialMediaUse2021 reported that although for some users (roughly one quarter) the effects of social media use on well-being were negative, for almost the same number of users they were positive, while for the rest the effects were neutral.
This finding is aligned with the Differential Susceptibility to Media Effects Model: 
Although there is substantial _variation_ of media effects for individual users, the _average_ effects reported in the literature are often small [@valkenburgDifferentialSusceptibilityMedia2013].

@@

What could be a minimally interesting, nontrivial effect?
Because this is a normative and ultimately philosophical question, there can never be a clear, single, or unanimous answer. 
<!-- In the end, it is a personal question. -->
However, it is still necessary and helpful to try to provide such a plausible benchmark. 
I therefore suggest the following SESOI for this research question:

@@

Thus, only when controlling for _all_ relevant confounders, can we correctly estimate causality without bias [@rohrerThinkingClearlyCorrelations2018].
Obviously, we can never be entirely sure to have included all confounders, which makes absolute statements regarding causality virtually impossible.
In addition, when determining the overall causal effect, we need to make sure _not_ to control for mediating variables [@rohrerThinkingClearlyCorrelations2018], for doing so would bias our assessment of the causal effect.
Complicating matters further, it is often unclear if a variable is a mediator or a confounder.[^collider]
However, despite all these caveats, when controlling for relevant variables (that aren't mediators), we can be much more certain that we measured causality correctly.
The aim should therefore be to collect as many varying and non-varying confounders as possible (which I believe is seldom done in our field), while knowing that absolute certainty regarding causality cannot be reached.

@@

Finally, one precondition of causality is temporal order. 
The cause needs to precede the effect.
Finding the right interval between cause and effect is crucial.
For example, if we want to understand the effect of alcohol consumption on driving performance, it makes a big difference if driving performance is measured one minute, one hour, one day, or one week after consumption.
<!-- Finding the right interval is difficult. -->
If variables are stable, longer intervals are needed; if they fluctuate, shorter intervals.
In the case of well-being, we need shorter intervals for **the more fluctuating** positive and negative affect and longer ones for **the more stable** life satisfaction [@dienlinImpactDigitalTechnology2020].
**Using social media can have instant effects on mood [@marcianoDynamicsAdolescentsSmartphone2022].**
**Effects on life satisfaction often take longer to manifest, for example because media use leads to actual changes in specific behaviors, which then affect life satisfaction [@dienlinDisplacementReinforcementReciprocity2017].**
Choosing the right interval is challenging, because especially short intervals are hard to implement in practice, often requiring advanced methods such as experience sampling (also known as in situ measurement or ambulant assessment) [@schnauber-stockmannMobileDevicesTools2020].
In this study, I hence analyze how using social media during the last week affected positive and negative affect during the same week.
In other words, if people during the last week engaged in more COVID-19 related social media use than they usually do, did they feel better or worse during that week than they usually do?
Regarding life satisfaction, I implemented a longer interval.
If people during the last week used COVID-19 related social media more than they usually do, were they at the end of the week more or less satisfied with their lives than they usually are?
I hence analyze if when a person changes their social media diet, are there (a) _simultaneous_ changes in their affect and (b) _subsequent_ changes in their life satisfaction?
**These relations will be controlled for varying confounders, which fosters a causal interpretation.**
Similar approaches were implemented by other studies [@johannesNoEffectDifferent2022; @scharkowHowSocialNetwork2020], and they are considered a best practice approach toward analyzing causality [@bellFixedRandomEffects2019].

@@

Because the data were analyzed post-hoc, no a-priori sample size planning on the basis of power analyses was conducted.
The sample is large, and it is hence well-equipped to reliably detect small effects.
<!-- , which is why no exact post hoc power analysis were conducted. -->
In addition, because such large samples easily generate significant _p_-values even for very small effects, it helps that the hypotheses were tested with a smallest effect size of interest-approach.
To this end, I adopted the interval testing approach as proposed by @dienesUsingBayesGet2014.
On the basis of the SESOI, I then defined a null region.
In what follows, I explain how I determined the SESOI and the null region.

@@

The factorial validity of the scales were tested with confirmatory factor analyses (CFA).
Because Mardia’s test showed that the assumption of multivariate normality was violated, I used the more robust Satorra-Bentler scaled and mean-adjusted test statistic (MLM) as estimator.
To avoid over-fitting, I tested the scales on more liberal fit criteria (CFI > .90, TLI > .90, RMSEA <. .10, SRMR < .10) [@klinePrinciplesPracticeStructural2016].
<!-- Finally, REWB-models cannot model latent variables. -->
<!-- To increase precision, I therefore exported factor scores from the CFAs for positive and negative affect. -->
Mean scores were used for positive and negative affect.
<!-- Respondents who answered less than 50% of all questions were removed. -->
**Missing responses were imputed using multiple imputation with predictive mean matching (five iterations, five data-sets), including categorical variables.**
**All variables were imputed except the media use measures, as they were not collected on each wave.**
**All variables included in the analyses presented here were used to impute missing data.**
**For the main analyses, results were pooled across all five data-sets.**

@@

**There was a small and statistically non-significant trend that reading COVID-19 related content slightly increased life satisfaction (_b_ = `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(conf.high) %>% round(2)`], _p_ = `r dat_fig_results_within %>% filter(dv == "Life satisfaction" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(p.value) %>% my_round("p")`).**
**At the same time, there was also a small and statistically non-significant trend that reading COVID-19 related content decreased positive affect (_b_ = `r dat_fig_results_within %>% filter(dv == "Positive affect" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(estimate) %>% round(2)` [95% CI `r dat_fig_results_within %>% filter(dv == "Positive affect" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(conf.low) %>% round(2)`, `r dat_fig_results_within %>% filter(dv == "Positive affect" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(conf.high) %>% round(2)`], _p_ = `r dat_fig_results_within %>% filter(dv == "Positive affect" & iv == "Reading" & type == "within" & Analysis == "Regular") %>% select(p.value) %>% my_round("p")`).**

@@

For between-person relations, no a-priori hypotheses were formulated.
Results showed that no relation crossed or was completely outside of the SESOI.
Four relations were statistically significant.
**Respondents who across all waves used social media more frequently than others to read about COVID-19 related posts reported slightly lower levels of positive affect than others (_b_ `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between" & Analysis == "Regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Reading" & type == "between" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`]).**
**Respondents who across all waves used social media more frequently than others to write COVID-19 related posts reported slightly higher levels of negative affect than others (_b_ `r dat_fig_results_between %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_between %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_between %>% filter(dv == "Negative affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`]).**
**At the same time, respondents who across all waves used social media more frequently than others to write COVID-19 related posts also reported slightly higher levels of positive affect (_b_ `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_between %>% filter(dv == "Positive affect" & iv == "Posting" & type == "between" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`]).**
**Finally, respondents who across all waves used YouTube more frequently than others also reported slightly higher levels of life satisfaction than others (_b_ `r dat_fig_results_between %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between" & Analysis == "Regular") %>% select(estimate) %>% my_round("b_txt")` [95% CI `r dat_fig_results_between %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between" & Analysis == "Regular") %>% select(conf.high) %>% my_round("b")`, `r dat_fig_results_between %>% filter(dv == "Life satisfaction" & iv == "YouTube" & type == "between" & Analysis == "Regular") %>% select(conf.low) %>% my_round("b")`]).**
<!-- However, note that the effect were still completely inside of the null region, hence not large enough to be considered practically relevant. -->

@@

**That said, several preliminary and subtle trends can be observed.**
**First, overall the results do suggest that effects of COVID-19 related social media use on well-being rather tend to take place in the negative as opposed to the positive spectrum.**
**For example, people who wrote more COVID-19 related posts than usual reported slightly lower levels of life satisfaction than usual.**
**Similarly, people who wrote more COVID-19 related posts than usual also reported slightly more negative affect.**
**When writing posts and comments on social media, people explicitly and more deeply engage with COVID-19 related content.**
**Unfortunately, in general online tonality is often more extreme, negative, or aggressive, which potentially affects their authors.**
**Note that I controlled for whether or not participants had a COVID-19 infection during a specific wave, which rules out the potential explanation that having an infection was the root cause of increased communication and reduced well-being.**

**The potential explanation that tonality might be a relevant factor at play here is also supported by the second trend.**
**People, who during the pandemic spent more time on Instagram than usual, also experienced less negative affect than usual.**
**Instagram is well-known for its positivity bias [@waterlooNormsOnlineExpressions2018].**
**Content is generally post positive, uplifting, and (self-)flattering.**
**It seems the often-criticized positivity bias on Instagram might have been somewhat beneficial in times of the pandemic.**
**The critique that the positivity bias necessarily leads to envy and negative feelings is one-sided, because positive content can also inspire and motivate users [@meierInstagramInspirationHow2020], which could be especially helpful in times of lockdown and home-office.**
**To provide a concrete example, Instagram was successfully used as an interactive communication channel for first year students to have a better start into their new degree, effectively complementing alternative learning platform tools [@yeTurningInformationDissipation2020].**

@@

**The results showed that it makes sense to analyze different communication types and communication channels, and that active and passive communication showed different results.**
**Liking and sharing content did not show any within-person effects. **
**Such rather low-key active behaviors do not seem to affect well-being at all.**
**Regarding passive use, reading COVID-19 related posts is more ambivalent; results showed some weak trends towards a positive effect on life satisfaction, but a negative effect on mood.**
**It might be that reading and informing oneself about COVID-19 on social media might be helpful in the long term, but more negative for short-term affect.**
**Finally, proactively engaging via writing posts, the most active form of communication analyzed here, showed only negative effects on well-being.**
**Together, the results support the findings from @valkenburgAssociationsActivePassive2022, who could not confirm the claim that active use is good, versus passive is bad.**
**Focusing on communication channels, YouTube seems to be more negative, whereas Instagram is likely more positive.**
**Again, these are only very small effects.**
**Future research might elaborate on these specific relations to probe their stability and relevance.**

**From a more political and societal perspective, the results imply that it can make sense to critically reflect upon COVID-19 related social media use.** 
**In terms of concrete recommendations, on average it might be slightly beneficial to post less actively about COVID-19 on social media and to spend less time on YouTube.**
**Potential positive effects, however, will for many users likely not be noticeable.**
**Results allow us to reject a positive effect: Writing more posts on social media will likely not increase well-being.**
**At all events, engaging in COVID 19-related social media use should, on average, not be a cause for concern.**

@@

The current study analyzed whether changes in media use were related to changes in well-being, while controlling for several potential confounders. 
Together, this allowed for an improved perspective on assessing causality.
**However, the opposite effect is also plausible, namely that well-being affects media use [@zillmannMoodManagementCommunication1988].**
**While controlling for potential confounders can support claims of causality, the procedure implemented here cannot prove causality.**
Causality necessitates temporal order, and the cause needs to precede the effect.
**The challenge is that** regarding media use, such effects often happen immediately or shortly after use, necessitating intervals in the hours, minutes, or even seconds.
In many cases only experience sampling studies asking users at the very moment can produce such knowledge.
However, even then we don't know for certain if we actually measured the right interval.
Effects depend on the intensity of use or the length of the interval.
To borrow the words from @rohrerTheseAreNot2021, there is no such thing as "the" effect of social media use on well-being.
Hence, to document how effects unfold, future research needs to employ different study designs probing different time lags.
In addition, more thought needs to be invested in what relevant stable and varying factors we should include as control variables, and I hope this study provides a first step into this direction.

Although I had already reduced the predefined SESOIs to be less conservative, they were potentially still too large.
Media use is only one aspect of several factors that simultaneously affect well-being.
Is it really realistic to expect that extremely changing only _one_ of these aspects should manifest in a detectable change in well-being?
Or would it make more sense to expect that thoroughly committing to say _two_ activities (e.g. regularly exercising _and_ establishing a reading habit) should then cause a detectable improvement in well-being?
Practically, this would imply a SESOI half as large as I have defined here, namely _b_ = |.15| for well-being and _b_ = |.075| for affect.
In the case of this study, however, reducing the SESOI would not even make a big difference, as also with these more liberal thresholds all but three effects would still be completely in the null region, and no effect would be outside of the null region.
However, at all events I encourage future research to start a thorough conversation on what effect sizes are considered meaningful and what not.
Again, with this study I hope to provide some first input and guidelines.
